Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics

### 2023-09-28 referee-replies.qmd

---
title: "Referee Replies"
format: pdf
---
## Notes
- emphasize that this is one model that doesn't capture a lot of stuff

"Spatial Heterogeneous" as definition for the kind of stock structure modeled here, goes from "panmictic" to "Spatial Heterogeneous", per the definitions in bosley et el. 2021. Acutally it can range all the way to metapopulation (set density dependence timing to local, shut down movement at one or more life stages)

## R1

Comments to the Author
This paper presents interesting, well-documented, and useful modeling software for spatial management of fisheries targeting multiple non-interacting species. I could see it becoming a go-to for both fisheries scientists and advanced fisheries courses. It should be published following minor revisions.

**We thank the reviewer for their helpful comments. **

My main suggestion for revision is to explain and emphasize the case-study results a bit more. Ultimately, the insights from the case studies will be one of the main things that convince readers to use the model. For example, what is the intuition behind the results in Figure 5A, showing target-fishing design being much more robust to range shifts than rate design? More generally, what are the most important things we should take away from the case studies?

**We have added a specific "Insights from Case Studies" section to the discussion to provide clearer insights from our results.**

Second, the authors should clarify that their model has both strategic and tactical uses, not just tactical. There are a lot of basic strategic questions that could be asked with this model. The case studies are arguably strategic as much or more so than they are tactical.

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

Minor comments:
Line 116: typo ‘th’.
**Corrected.**

Line 169: This seems like a potentially strong assumption that could artificially eliminate weak-stock-strong-stock tradeoffs. This should be discussed and justified.

**This is a very good point, particularly as 0% discard mortality is clearly the exception rather than the rule. We agree that this is too strong of assumption to use in many cases, and are working to implement both a discard rate and a discard mortality function into the model. However, as seasonal management is not a key feature of the model at this time, and is not used in our case study, we have elected to remove the mention of seasonal closures for now from the manuscript, though we are pursuing the addition of this feature in the package itself.**

XX Probably not going to do this yet since it is a pain in the ass XX. 
**To address this concern, we have modified the model such that "closed" seasons are now implemented by a multiplier on fishing mortality that can range from 0 to 1. When the multiplier is set to 0, the closed season is 100% effective (i.e. no discard mortality). Setting the multiplier to 0.5 instead reduces F by 50%, reflecting a 50% discard mortality rate**

Line 280: missing period.

**Corrected.**

Line 283: that is a very nice model feature.

**We are glad the reviewer finds this feature helpful, many thanks to the folks at FishBase and Thorson's prior work with FishLife for making it possible.**

Line 301: Is omitting advection only a significant limitation for drifting species like plankton or jellyfish, or is it also a limitation for large pelagics that concentrate at SST fronts?

**Good question. A strength of the CTMC approach is that, given sufficient data, one can separate out the components of movement into advection, taxis, and diffusion. Having this knowledge is certainly interesting, but from a simulation perspective is less critical. One could simulate the dynamics of a large pelagic concentrating at an SST front using simply the taxis matrix in `marlin`; while the SST fronts themselves are clearly a partial product of advective forces, so long as we know where they are, we can just pass their location to marlin and the fish will go there. What would be missing there might be some timing dynamics whereby a fish might take longer to get to ideal habitat when fighting against a current than with it. The biggest limitation of omitting the advection matrix at this point is the potential to overlook cases where a current is strong enough to essentially serve as a barrier preventing an organism from reaching preferred habitat that it otherwise would go to under the taxis + diffusion model. We have clarified this in the text of the discussion, and pointed out the value of research into the importance of advective forces as well. **

Line 473: Typo: “Do” -> “Due”.
**Corrected.**

##R2

FaF-23-Apr-OA-149 Review 02 Aug 2023
Summary
This study presents a bio-economic modeling tool aimed to provide a tractable, efficient, tactical management simulation framework for marine fisheries. It seeks to complement earlier, similar, models which are highly complex, bespoke, or overly-generalized. Case study applications to a coral reef and pelagic ecosystem is presented. The study conclusions does not evaluate spatial management paradigms per se, but rather illustrate the value of “sufficiently realistic” biodynamic models for policy evaluation.

This model and paper provide the following novel contributions:
1) the modeling framework allows users to test either eliminating or displacing fishing effort given the placement of an MPA; ignoring these alternative possibilities was the cause of much scandal and scientific misinformation in recent MPA analyses,

**It certainly has been, we are glad the reviewer finds this feature useful**

2) a wise avoidance of the calculation of equilibrium-based quantities for spatio-temporal models, precluding scientists’ conclusion that certain policies do or do not meet management targets (as these are not definable under such conditions), and

**Though dynamic and spatially explicit reference points are certainly an area of interesting research going forward.**

3) an explicit manner of investigating MPA effectiveness in the context of range shifts, which are important for climate-enhanced fishery management.

I recommend publication with minor revision.

Major Comments
This is a well-written paper, and I appreciated that the methods are clear and high-level, with minimal equations. There are two central things that need to be addressed in the document, which would bolster and clarify how marlin fits into the spatial modelling world, and how we can interpret the results. I believe these can be readdressed through a fairly simple simulation exercise and the addition of a few paragraphs to the methods & discussion.

Most importantly, the reader wants to know whether marlin is providing insights consistent with our understanding of how population and fishery dynamics behave in space, and whether the reader has enough information to make it worth testing the model on their fishery. If the aim is for this model to be used by global communities with little data or information about their fishery, please discuss if/how the model can be simplified to provide insights useful in the absence of key processes (i.e., selectivity).

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

How dependent is the simulation accuracy on full confidence in the inputs? I am not requesting a full misspecification analysis. Rather, please illustrate or discuss how the results of the coral reef case study, for example, might change if we are incorrect that Fleet Two has dome-shaped selectivity. We know this to be problematic in spatial models. How do your findings compare to earlier examinations of mis- specified selectivity in spatial context, e.g. works by Aaron Berger and Lynn Waterhouse?

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

XX I still need to integrate some more of these XX s

Bosley, Katelyn M., Amy M. Schueller, Daniel R. Goethel, Dana H. Hanselman, Kari H. Fenske, Aaron M. Berger, Jonathan J. Deroba, and Brian J. Langseth. “Finding the Perfect Mismatch: Evaluating Misspecification of Population Structure within Spatially Explicit Integrated Population Models.” Fish and Fisheries 23, no. 2 (2022): 294–315. https://doi.org/10.1111/faf.12616.

Berger, Aaron M., Daniel R. Goethel, Patrick D. Lynch, Terrance Quinn, Sophie Mormede, Jeremy McKenzie, and Alistair Dunn. “Space Oddity: The Mission for Spatial Integration.” Canadian Journal of Fisheries and Aquatic Sciences 74, no. 11 (November 2017): 1698–1716. https://doi.org/10.1139/cjfas-2017-0150.

Berger, Aaron M., Michael L. Jones, Yingming Zhao, and James R. Bence. “Accounting for Spatial Population Structure at Scales Relevant to Life History Improves Stock Assessment: The Case for Lake Erie Walleye Sander Vitreus.” Fisheries Research 115–116 (March 1, 2012): 44–59. https://doi.org/10.1016/j.fishres.2011.11.006.

Berger, Aaron M, Jonathan J Deroba, Katelyn M Bosley, Daniel R Goethel, Brian J Langseth, Amy M Schueller, and Dana H Hanselman. “Incoherent Dimensionality in Fisheries Management: Consequences of Misaligned Stock Assessment and Population Boundaries.” ICES Journal of Marine Science 78, no. 1 (February 1, 2021): 155–71. https://doi.org/10.1093/icesjms/fsaa203.

Waterhouse, Lynn, David B. Sampson, Mark Maunder, and Brice X. Semmens. “Using Areas-as-Fleets Selectivity to Model Spatial Fishing: Asymptotic Curves Are Unlikely under Equilibrium Conditions.” Fisheries Research, SI: Selectivity, 158 (October 1, 2014): 15–25. https://doi.org/10.1016/j.fishres.2014.01.009.

- Add discussion to the effect of "marlin has a lot of options. A recommended use case would be to try and narrow down things that are "known", and then focus sensitivity runs on outcomes that are uncertain and likely to be important. For example, testing the sensitivity of outcomes to misspecified selectivity curves

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

The value of this contribution would be stronger if you directly compared the findings from your case studies with similar works. For example, there have been several studies (empirical and simulation- based) examining MPAs and coral reef ecosystems. Several have found confounding effects of MPA closures, often caused by inter-species interactions, which I recognize are not modeled here. Does marlin provide similar insights on less information (i.e. without predator dynamics)?

**case study section literature time here. Really struggling with this one. Thoughts on how to address?**

Of secondary importance, I am not sure I agree that marlin follows the MICE paradigm. Much of the model specification (I’m thinking movement especially) is highly granular even for well-studied fisheries.

**We have removed the reference to MICE models per say..**

It is common for tactical applications to model irregular polygons of distinct biological and/or political attributes, connected by a user-input matrix of movement values. Confirm whether the current setup requires individual patches to share biological information and the harvest strategies among them. Discuss how spatial heterogeneity in population demography can impact model results.

**This is a very good point; spatio-temporal variation in population demography (and regulation) can be extremely important and is not captured here. `marlin` assumes that while habitat can vary in space and time, population processes are constant within a species. That being said,  users could treat different morphs of a population as separate species with different habitats. For example, if an invertebrate species grows at a different rate in warmer coastal waters than cooler deeper waters, these could simply be modeled as two separate species with different growth parameters. However, the model would not then allow for interbreeding of those morphs..xx** Cite @waterhouse2014 here, need to find where to cite and add to paper XX

see "finding the perfect mismatch bosley paper" for discussions around frequency of spatial complexity paired with frequency of ignoring it

Minor Comments
I recognize the authors didn’t want to detail every aspect of the model specifications, but these are things the reader might want to know from the outset:

L106 “10 x 10 patch system” – units? what is a “patch” here? (related to point above about what info is shared within a patch)

**Users on standard computers should be able to simulate one age-structured population distributed across spatial surface represented by a ten by ten grid of cells over 20 years in fractions of a second**

L157 you mention size limits as a potential harvest control – but don’t say that size is also represented in the pop dynamics until much later

**We have now clarified in the opening description that the model also tracks lengths at age**.

Model Summary – somewhere early in here, confirm that there is no estimation/conditioning capacity within marlin (to discourage folks from thinking they might be able to run assessments within it, for example). Add a sentence or two near the beginning describing what the user must provide to the model, and state that meta-analyses like FishLife will be consulted in their absence.

**We have added a few sentences to the model summary section describing the minimum inputs needed, and making the important point that there is no estimation feature built into marlin**

L232 Related to above: I see the description of the tuning settings. When total SSB is specified in the context of MPA, how is Q adjusted? Does this step interact with the option to either displace or turn off effort outside of an MPA?

**We have clarified in the text that "`marlin` allows users to specify a target unfinished spawning stock biomass ($SSB0$), which will be achieved by tuning the total unfished recruitment (R0), given the remaining life history parameters and independent of any characteristics of the fishing fleets.". We have also clarified in the "tuning fishing fleet parameters" section that all tuning is done conditional on the underlying population dynamics.**

L259 what sort of time scale is needed for this approximation to be true?

**We have clarified in the text that "The timeline required for the fleets to reach an equilibrium distribution will vary and depend on factors such as the population dynamics of the species in question (e.g. movement, habitat heterogeneity, growth rates, density dependence) and the number and degree of competition across fleets. Users should explore different simulation times to ensure that any results they wish to use are not simply a reflection of the fleet dynamics fluctuating on their way to an equilibrium condition. "**

L310 “that the model is running under” – rephrase

**Clarified**.

L363 be consistent about capitalization of SSB

**Fixed**

L447 too many quotes in this section

**Removed**

L493 quarterly timestep? Only one year (four seasons total)?

**We have clarified that the simulation runs for 100 years in time steps of quarter years (i.e. four seasons per year)**

Figure 3 I’d prefer this to have the same colorscale as Figure 1; do a panel for each condition, don’t overlay them.

**We have changed Figure 3 to reflect the same format as Figure 1, and agree it is much improved, thank you!**

L534 Do they actually mask them, or do interesting dynamics underlie these findings? Add a sentence here what users should do with this info (i.e. be sure to look at the tradeoffs one by one – is there functionality in marlin to do so, such as automated plots or summary tables?)

**Good point, we clarify that "These macro-level changes in total yield and conservation are the result of interesting internal dynamics that `marlin` allows users to explore through output tables at the level of individual species and fleets." We do not specify specific functions within the marlin package here as the details of those may evolve over time.**

FIG 4. The results state three MPA strategies were used: spawning ground, spawning aggregation, and target fishing, but I only see two here.

**We have fixed this typo (we had accidentally refered to the same policy by two different names).**

FIG 4C This figure is confusing; I’m not clear what the skinny short lines are. The caption doesn’t clarify which color of label corresponds to which % change in yield; instead just use a colored point for these values.

**We have modified Figure 4C and we hope the reviewer finds it clearer now. Much of the visual clutter of Fig 4C was due to the long tail to the right of high SSB increases and catch decreases that occured once MPA size approaches 100%. To help illustrate the more interesting points of the results, we have limited the figure to MPA sizes <= 65%, and removed the text labels for MPA sie and replaced them with colored points to indicate MPA size along each of the MPA design strategy curves.**

### 2023-09-28 03_marlin_paper.qmd
getting somewhere

---
format:
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
    keep-tex: true
    fig-format: png
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: apa-6th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.0"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{bm}
---

@@

@@

The front end of the software package accompanying this paper,
[`marlin`](https://danovando.github.io/marlin/), is written in R
[@rcoreteam2021] to facilitate use, while the underlying population
model is written in C++, integrated through the `Rcpp` package
[@eddelbuettel2018], for speed. Users on standard computers should be
able to simulate one age-structured population distributed across a 10
by 10 patch system for 20 years in fractions of a second.

@@

We define a few commonly used terms here. *Yields* refers to the volume
of catch from fishing activity. *Spawning stock biomass* refers to the
total biomass of reproductively mature fish in the water, a function of
numbers, weight, fecundity, and sexual maturity at age. We measure the
size of the population by th ratio of the *spawning stock biomass* (SSB)
in a given time step relative to *unfished spawning stock biomass*
(SSB0), SSB/SSB0. So, a SSB/SSB0 value of 1 means that the population is
unfished, a SSB/SSB0 value of 0 means the population is extinct.

@@

`marlin` simulates the dynamics of one or more species, currently best
representing fish-like species, using age-structured population
dynamics. Each time step, species move throughout the simulated seascape
using both diffusion and "taxis" (active movement towards preferred
habitat), experience natural and potentially fishing mortality, and if
in a spawning season spawn using one of the possible forms of density
dependence implemented in the model, with the potential to include
auto-correlated stochastic deviations in the amount of offspring
produced, generally called "recruitment deviates".

@@

Users can use the model to explore general properties of spatially
explicit social-ecological systems, or can parameterize it to explore
key uncertainties of more tactical applications. As an example of the
later case, we provide two case studies here designed to reflect the
kinds of questions stakeholders might face if designing an MPA network
for a coastal coral reef or a pelagic system.

@@

Users can tune their fleets in one of two ways. First, they can specify
a target exploitation rate *u* for each species in their simulation.
Taking all the other parameters of the model as given, `marlin` then
adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and
species *s* such that the equilibrium exploitation rate for each species
matches the desired level. Second, they can specify a target total
spawning stock biomass under fishing divided by total unfished spawning
stock biomass, and the model will adjust the catchability coefficients
$q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium
ratio of fished to unfished spawning biomass for each species matches
the desired level.

@@

As such we explored a series of "next best" fleet distribution
algorithms. While not the IFD in any individual time step, over time
they start to approximate the IFD, as each assumes that fleets base
their decisions for the current time step on the outcomes in the prior
timestep, meaning that the impacts of the actions of other fishing
fleets are somewhat internalized. The four possible fleet distribution
algorithms are

@@

Revenue based spatial distribution is not likely to be very realistic;
in general we would expect fishing fleets to respond to profits on some
core level. However, due to the complexity of parameterizing cost
functions, fleet dynamics are often evaluated based on yield or revenue
alone, and so we include those scenarios here to allow users to evaluate
the potential implications of this choice. The decision on whether to
allocate the fleet based on absolute or relative (per unit effort)
metrics is more complex. When effort represents the actions of separate
and individual fishing actors (e.g. independent fishing vessels), a
per-unit-effort strategy, in which fishers distribute themselves based
on the expected catch of their individual efforts, may be more realistic
[@hilborn1987]. Conversely, a system defined by a sole owner seeking to
maximize total profits might be better represented by a fleet model
based on total profits

@@

We provide a brief overview of the the general CTMC method here (see
@thorson2021a for a detailed description). Under this framework,
movement of individuals from each patch to each other patch in the
system in a given timestep *t* for life stage *a* of species *s* is
defined by a movement matrix $\pmb{M}_{t,s,a}$. $\pmb{M}_{t,s,a}$ is
calculated as a function of diffusion $\pmb{D}$ and taxis $\pmb{\tau}$
matrices scaled by the width of the time step $\Delta_{t}$ and the
length of the edge of each patch $\Delta_d$ that the model is running
under. This parameterization allows users to set the effective area of
the spatial domain through two avenues; the number of patches, which
effectively scales the resolution of the model, and the area of each
patch, which scales the spatial extent of the simulation.

@@

For the population model, numbers *n* at time *t* for age *a* are a
function of growth, death, and recruitment

$$
n_{t,p,s,a} = \begin{cases}
      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
  \end{cases}
$$ {#eq-pop}

where *BH* is the Beverton-Holt recruitment function [@beverton1957] and
*ssb* is spawning-stock-biomass. Per convention, the model allows for a
"plus group", wherein rather than tracking numbers of every possible
age, individuals greater than or equal to a given maximum age are
grouped together.

Spawning stock biomass *ssb* is calculated by converting age to mean
length, calculating weight at age, maturity at age, and then calculating
spawning stock biomass as the sum of spawning potential at age in a
given time step, taking into account the potential for hyperallometry in
the manner of @marshall2021. Age is converted to length through the von
Bertalanffy growth equation given parameters asymptotic length
($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$)
assuming log-normally distributed variation *u* in the length at age
with CV $\sigma_s$.

@@

Biomass *b* at age is then given by the weight at length equation
governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$
that controls the rate at which volume scales with length

$$
  b_{a,s} = _s \times l_{a,s}^{wb_s}
$$ {#eq-weight}

@@

$$
  ssb_{t,p,s} = \sum_{a=1}^{N_a}w_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
$$ {#eq-ssb}

@@

Recruitment (i.e. the number of age 1 individuals entering the
population) follows Beverton-Holt dynamics parameterized around
steepness (*h*) with log-normally distributed recruitment deviates
$\epsilon$. When steepness is one recruitment is independent of spawning
biomass. As steepness approaches 0.2 recruitment becomes a linear
function of spawning biomass. We allow for five variants in the timing
of density dependent recruitment, building off of @babcock2011 :

@@

$$
\begin{aligned}
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
}
\end{aligned}
$$ {#eq-dd2}

@@

$$
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right) \times \epsilon_{t,s}
$$ {#eq-dd1}

@@

$$
  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right)\times \pmb{d^l_s} \times \epsilon_{t,s}
$$ {#eq-dd3}

@@

$$
  larv_{t,p,s} = ssb_{t-1,p,s} \times\pmb{d^l_s}
$$ {#eq-larvmove}

$$
n_{t,a = 1,p,s} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{larv_{t,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{larv_{t,p,s}}}\right) \times \epsilon_{t,s}
$$ {#eq-dd4}

@@

$$
\begin{aligned}
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ \frac{ssb_{t-1,p,s}}{\sum_{p=1}^Pssb_{t-1,p,s}} \times \epsilon_{t,s}
\end{aligned}
$$ {#eq-dd5}

@@

3.  *Spawning Aggregation*: MPAs are centered on the location of a known
    spawning aggregation

To our knowledge, MPAs are never randomly placed; their locations are
decided based on stakeholder negotiation given available information and
objectives. In theory, the design of MPA networks can be optimized
through the use of a modeling framework, and depending on the validity
of the model, this process is likely to produce better outcomes than
manually-designed strategies [@rassweiler2012; @rassweiler2014].
However, designing optimal MPA networks becomes increasingly difficult
as the range of objectives and the complexity of the model increase.
Therefore, while `marlin` provides some support for
[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
we focus here on the design and performance of these more "rule of
thumb" design strategies that may be more accessible to a wider range of
users. We allow all MPAs to be designed in a "mosaic" fashion in these
examples, but users can easily extend the analysis to compare outcomes
between "contiguous" (MPA is made up of one continuously connected
block) and "mosaic" (MPAs can be separated in space) MPA designs
[@pons2022] by developing and passing their own MPA design algorithms to
the model.

@@

>>>>

---
format:
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
    keep-tex: true
    fig-format: png
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: apa-6th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.2"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{bm}
editor: 
  markdown: 
    wrap: 80
---

@@

## Need to cite..

@bosley2022

@berger2017

@@

The front end of the software package accompanying this paper,
[`marlin`](https://danovando.github.io/marlin/), is written in R
[@rcoreteam2021] to facilitate use, while the underlying population model is
written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for
speed. Users on standard computers should be able to simulate one age-structured
population distributed across two-dimensional spatial surface represented by a
ten by ten grid of cells over 20 years in fractions of a second.

@@

We define a few commonly used terms here. *Yields* refers to the volume of catch
from fishing activity. *Spawning stock biomass* refers to the total biomass of
reproductively mature fish in the water, a function of numbers, weight,
fecundity, and sexual maturity at age. We measure the size of the population by
the ratio of the *spawning stock biomass* (SSB) in a given time step relative to
*unfished spawning stock biomass* (SSB0), SSB/SSB0. So, a SSB/SSB0 value of 1
means that the population is unfished, a SSB/SSB0 value of 0 means the
population is extinct.

@@

`marlin` simulates the dynamics of one or more species, currently best
representing fish-like species, using age-structured population dynamics. Ages
are then converted to lengths using the von Bertalanffy growth equation with
log-normally distributed variation in the length at age. Each time step, species
move throughout the simulated seascape using both diffusion and "taxis" (active
movement towards preferred habitat), experience natural and potentially fishing
mortality, and if in a spawning season spawn using one of the possible forms of
density dependence implemented in the model, with the potential to include
auto-correlated stochastic deviations in the amount of offspring produced,
generally called "recruitment deviates".

@@

Users can use the model to explore general properties of spatially explicit
social-ecological systems, or can parameterize it to explore key uncertainties
of more tactical applications. It is very important note that the parameters of
this model cannot be fit to data directly within the `marlin` package; users
must set model parameters themselves based on externally available data and
their best judgement. On the biological side, at a minimum users must for each
species they wish to simulate supply the common or scientific name of the
species in question, a measure of the level of fishing intensity the species is
experiencing at the start of the simulation, and the diffusion rates for adult
and larvae. Given a common or scientific name, the model will then input default
life history parameters based on FishLife [internet connection required,
@thorson2020], though users are encouraged to check these values and supply
their own life history parameter values when possible. On the fleet side, users
must for each fleet they wish to simulate at a minimum specify the selectivity
curves for each species caught by the fleet, as well relative price per unit
weights greater than 0 for any target species. Species caught by but not
targeted by the fleet can be represented by prices of 0 (not targeted) or below
0 (actively avoided). We provide two case studies here designed to show users
how the `marlin` model can be used to address the kinds of questions
stakeholders might face if designing an MPA network for a coastal coral reef or
a pelagic system.

@@

Users can tune their fleets in one of two ways, conditional on the underlying
population dynamics of the species in question. First, they can specify a target
exploitation rate *u* for each species in their simulation. Taking all the other
parameters of the model as given, `marlin` then adjusts the catchability
coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the
equilibrium exploitation rate for each species matches the desired level.
Second, they can specify a target total spawning stock biomass under fishing
divided by total unfished spawning stock biomass, and the model will adjust the
catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that
the equilibrium ratio of fished to unfished spawning biomass for each species
matches the desired level.

@@

As such we explored a series of "next best" fleet distribution algorithms. While
not the IFD in any individual time step, over time they start to approximate the
IFD, as each assumes that fleets base their decisions for the current time step
on the outcomes in the prior timestep, meaning that the impacts of the actions
of other fishing fleets are somewhat internalized. The timeline required for the
fleets to reach an equilibrium distribution will vary and depend on factors such
as the population dynamics of the species in question (e.g. movement, habitat
heterogeneity, growth rates, density dependence) and the number and degree of
competition across fleets. Users should explore different simulation times to
ensure that any results they wish to use are not simply a reflection of the
fleet dynamics fluctuating on their way to an equilibrium condition.

The four possible fleet distribution algorithms are

@@

Revenue based spatial distribution is not likely to be very realistic; in
general we would expect fishing fleets to respond to profits on some core level.
However, due to the complexity of parameterizing cost functions, fleet dynamics
are often evaluated based on yield or revenue alone, and so we include those
scenarios here to allow users to evaluate the potential implications of this
choice. The decision on whether to allocate the fleet based on absolute or
relative (per unit effort) metrics is more complex. When effort represents the
actions of separate and individual fishing actors (e.g. independent fishing
vessels), a per-unit-effort strategy, in which fishers distribute themselves
based on the expected catch of their individual efforts, may be more realistic
[@hilborn1987]. Conversely, a system defined by a sole owner seeking to maximize
total profits might be better represented by a fleet model based on total
profits.

@@

We provide a brief overview of the the general CTMC method here (see
@thorson2021a for a detailed description). Under this framework, movement of
individuals from each patch to each other patch in the system in a given
timestep *t* for life stage *a* of species *s* is defined by a movement matrix
$\pmb{M}_{t,s,a}$. $\pmb{M}_{t,s,a}$ is calculated as a function of diffusion
$\pmb{D}$ and taxis $\pmb{\tau}$ matrices scaled by the width of the time step
(e.g. one year) $\Delta_{t}$ and the length of the edge of each patch (e.g. one
kilometer) $\Delta_d$ specified by the user. This parameterization allows users
to set the effective area of the spatial domain through two avenues; the number
of patches, which effectively scales the resolution of the model, and the area
of each patch, which scales the spatial extent of the simulation.

@@

For the population model, numbers *N* at time *t* for age *a* are a function of
growth, death, and recruitment

$$
N_{t,p,s,a} = \begin{cases}
      = BH(SSB_{t-1,p,s,a}) & \text{if $a = 1$}\\
     = N_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
     =  N_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + N_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
  \end{cases}
$$ {#eq-pop}

where *BH* is the Beverton-Holt recruitment function [@beverton1957] and *SSB*
is spawning-stock-biomass. Per convention, the model allows for a "plus group",
wherein rather than tracking numbers of every possible age, individuals greater
than or equal to a given maximum age are grouped together.

Spawning stock biomass *SSB* is calculated by converting age to mean length,
calculating weight at age, maturity at age, and then calculating spawning stock
biomass as the sum of spawning potential at age in a given time step, taking
into account the potential for hyperallometry in the manner of @marshall2021.
Age is converted to length through the von Bertalanffy growth equation given
parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at
length zero ($a0$) assuming log-normally distributed variation *u* in the length
at age with CV $\sigma_s$.

@@

Biomass *B* at age is then given by the weight at length equation governed by a
scaling coefficient $\Omega_s$ and an exponent $\Phi_s$ that controls the rate
at which volume scales with length

$$
  B_{a,s} = _s \times l_{a,s}^{wb_s}
$$ {#eq-weight}

@@

$$
  SSB_{t,p,s} = \sum_{a=1}^{N_a}w_{a,s,t}^{\gamma_s} mat_{t,a} N_{t,p,s,a}
$$ {#eq-ssb}

@@

Recruitment (i.e. the number of age 1 individuals entering the population)
follows Beverton-Holt dynamics parameterized around steepness (*h*) with
log-normally distributed recruitment deviates $\epsilon$. When steepness is one
recruitment is independent of spawning biomass. As steepness approaches 0.2
recruitment becomes a linear function of spawning biomass. `marlin` allows users
to specify a target unfinished spawning stock biomass ($SSB0$), which will be
achieved by tuning the total unfished recruitment ($r0$), given the remaining
life history parameters and independent of any characteristics of the fishing
fleets.

We allow for five variants in the timing of density dependent recruitment,
building off of @babcock2011 :

@@

$$
\begin{aligned}
N_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{SSB_{t-1,p,s}}}{0.2\times{\sum_{p=1}^PSSB0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^PSSB_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
}
\end{aligned}
$$ {#eq-dd2}

@@

$$
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{SSB_{t-1,p,s}}}{0.2\times{SSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{SSB_{t-1,p,s}}}\right) \times \epsilon_{t,s}
$$ {#eq-dd1}

@@

$$
  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{SSB_{t-1,p,s}}}{0.2\times{SSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{SSB_{t-1,p,s}}}\right)\times \pmb{d^l_s} \times \epsilon_{t,s}
$$ {#eq-dd3}

@@

$$
  larv_{t,p,s} = SSB_{t-1,p,s} \times\pmb{d^l_s}
$$ {#eq-larvmove}

$$
n_{t,a = 1,p,s} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{larv_{t,p,s}}}{0.2\times{SSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{larv_{t,p,s}}}\right) \times \epsilon_{t,s}
$$ {#eq-dd4}

@@

$$
\begin{aligned}
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{SSB_{t-1,p,s}}}{0.2\times{\sum_{p=1}^PSSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^PSSB_{t-1,p,s}}}\right) \times \\ \frac{SSB_{t-1,p,s}}{\sum_{p=1}^PSSB_{t-1,p,s}} \times \epsilon_{t,s}
\end{aligned}
$$ {#eq-dd5}

@@

3.  *Spawning Ground*: MPAs are centered on the grounds of a known spawning
    aggregation

To our knowledge, MPAs are never randomly placed; their locations are decided
based on stakeholder negotiation given available information and objectives. In
theory, the design of MPA networks can be optimized through the use of a
modeling framework, and depending on the validity of the model, this process is
likely to produce better outcomes than manually-designed strategies
[@rassweiler2012; @rassweiler2014]. However, designing optimal MPA networks
becomes increasingly difficult as the range of objectives and the complexity of
the model increase. Therefore, while `marlin` provides some support for
[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
we focus here on the design and performance of these more rule of thumb design
strategies that may be more accessible to a wider range of users. We allow all
MPAs to be designed in a mosaic fashion in these examples, but users can easily
extend the analysis to compare outcomes between contiguous (MPA is made up of
one continuously connected block) and mosaic (MPAs can be separated in space)
MPA designs [@pons2022] by developing and passing their own MPA design
algorithms to the model.

@@

### General Recommendations for Using `marlin`

Marine social-ecological systems are highly complex. The model presented here is
designed to help users explore the impacts of different variables on policy
outcomes. But, that freedom comes at a cost, namely that users have a large
number of options at their disposal that they must decide on. Our recommended
strategy for dealing with this is for users to try and narrow down a list of
parameters that they feel are sufficiently "known" and another list of
parameters that are highly uncertain and / or contentious which the users feel
may impact results.
